The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-05-07 writeup.Rmd
Masoud completed his tasks for draft 1

```{r load_packages, include = FALSE}
library(grid)
library(gtable)
library(papaja)
library(tidyverse)
library(magrittr)
library(readr)
library(png)
library(jpeg)
library(lme4)
library(ggthemes)
library(forcats)
library(brms)
library(DescTools)
library(binom)
theme_set(theme_bw(18))
```

@@

Let's examine an example. Assume three response options (wrong, neither, right). Assume further that a third of participants each gave each of the three responses, i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are for instance interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are.

In the experiment reported in the following section, we presented participants with exactly this setup. Different groups of participants were presented with different numbers of response options. We categorized their responses according to the Weak and the Strong link and tested whether number of response options and categorization scheme leads to different conclusions about implicature rates.

@@

@@

```{r participants}
participants_info <-
  data %>%
  group_by(response_type) %>%
  summarise(count = n()/24)
```

200 participants were recruited using Amazon Mechanical Turk (binary=`r participants_info$count[1]`, ternary`r participants_info$count[4]`, quaternary=`r participants_info$count[2]`, quinary=`r participants_info$count[3]`). No participant was excluded from the final analysis.

### Procedure

The study was administered online through Amazon Mechanical Turk. Participants were introduced to a set of cards with pictures of one or two animals (Figure \@ref(fig:stimuli)). They were told that a blindfolded fictional character called Bob is going to guess what animals are on the card. On each trial, participants saw a card as well as a sentence representing Bob's guess. For example, they saw a card with a cat on it and read the sentence "There is a cat on the card." The study ended after 24 trials. At the end participants optionally provided demographic information. We also asked participants if they had any prior training in logic. You can access and view on the study's [online repository](https://github.com/thegricean/si-paradigms/).

### Design and Materials

@@

The study had two main manipulaitons within participants: the type of card and the type of guess. There were two types of cards. Cards with only one animal on them and cards with two animals. Animals were chosen from the following set: cat, dog, and elephant  There were three types of guesses: simple (e.g. *There is a cat*), conjunctive (e.g. *There is a cat and a dog*), and disjunctive (e.g. *There is a cat or a dog*). In each trial, the animal labels used in the guess and the animal images on the card may have no overlap (e.g. Image: cat, Guess: *There is an elephant*), a partial overlap (e.g. Image: cat, Guess: *There is a cat or a dog*), or a total overlap (e.g. Image: cat and dog, Guess: *There is a cat or a dog*). Crossing the number of animals on the card, the type of guess, and the overlap between the guess and the card results in 12 different possible trial types. We chose 8 trial types (Figure \@ref(fig:trials)), balancing the number of one-animal vs. two-animal cards, simple vs. connective guesses, and expected true vs. false trials. Three trials were randomly selected from each of the 8 trial-types, for a total of 24 trials. The order of these 24 trials was randomized as well.

@@

The study also had a between participant manipulation of the number of response options in the forced choice task. Participants were randomly assigned to one of four different conditions. The conditions differed with respect to the number of response options: binary (wrong vs. right), ternary (wrong, neither, right), quaternary (wrong, kinda wrong, kinda right, right), and quinary (wrong, kinda wrong, neither, kinda right, right). We wanted to see if the number of response options in the forced choice task would affect our estimate of the task's "implicature rate".

```{r trials, fig.asp=0.35, fig.cap = "Trial types represented by example cards and guesses."}
trials_img <- jpeg::readJPEG("experiments/figs/trialTypes.jpg")
grid::grid.raster(trials_img)
```

@@

<!--We used `r cite_r("r-references.bib")` for all our analyses.-->

@@

binary_summary_XY_and <-
  binary_summary %>%
  filter(card_type=="X", guess_type=="XandY")
binary_summary_XY_and_confint <-
  binary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
binary_summary_XY_and <- binary_summary_XY_and %>% full_join(binary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r binaryPlot, fig.height=3.5, fig.cap = "Proportion responses for the two-alternative (binary) forced choice judgments in the guessing game."}
binary_plot<-
  binary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_grid(card_type~guess_type) +
  labs(x="Response", y="Proportion of response")+
  theme_few() +
  theme(text = element_text(size=12)) +
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE) +
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

Here we present the proportion that participants chose the response options in each of the 8 trial types of the binary, ternary, quaternary, and quinary tasks. Figure \@ref(fig:binaryPlot) shows the proportion of "right" and "wrong" responses in the binary task. Starting from the leftmost column, participants considered a guess "wrong" if the guessed animal was not on the card. Moving to the second column, participants considered a guess "right" if the animal on the card was mentioned. However, if only one of the two animals on the card was mentioned (exhaustive trials), `r round(binaryExh_wrong$proportion*100)`% of the times participants considered the guess "wrong". Moving to the third column, if a conjunction of animals was guessed while only one animal was on the card, participants considered the guess to be "wrong". If a conjunction of animals was guessed and both animals were present on the card, all participants considered the guess to be "right" as expected. Moving to the forth column, if a disjunction of animals was guessed and only one of the animals was on the card, participants considered the guess to be "right" almost all the time. However, if both animals were present (scalar trials), `r round(binaryScalar_wrong$proportion*100)`% of the times participants considered the guess to be "wrong".

@@

ternary_summary_XY_and <-
  ternary_summary %>%
  filter(card_type=="X", guess_type=="XandY")
ternary_summary_XY_and_confint <-
  ternary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
ternary_summary_XY_and <- ternary_summary_XY_and %>% full_join(ternary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r ternaryPlot, fig.height=3.5, fig.cap = "Proportion responses for the three-alternative (ternary) forced choice judgments in the guessing game."}
ternary_plot<-
  ternary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_grid(card_type~guess_type) +
  labs(x="Response", y="Proportion of response")+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = c("red4","grey", "springgreen3"), guide=FALSE) +
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

Figure \@ref(fig:ternaryPlot) shows the proportion of "right", "neither", and "wrong" responses in the ternary task. Similar to the binary task, participants considered a guess wrong when the mentioned animal was not on the card. They considered the guess "right" when the mentioned animal was on the card. However, in exhaustive trials when the fictional character only guessed one of the two animals on the card, participants considered the guess "wrong" `r round(ternaryExh_wrong$proportion*100)`% of the time and neither wrong nor right `r round(ternaryExh_neither$proportion*100)`% of the time. If a conjunction of animals was guessed and only one animal was present on the card, participants considered the guess "wrong". As expected, when a conjunction was used and both animals were present, participants considered the guess "right". Similarly, participants considerd the guess "right" when a disjunction was used and only one of the animals was on the card. However, in scalar trials that both animals were on the card and a disjunction was gueassed, participants judged the guess "wrong" `r round(ternaryScalar_wrong$proportion*100)`% of the time and "neither" `r round(ternaryScalar_neither$proportion*100)`% of the time.

@@

quaternary_summary_XY_and <-
  quaternary_summary %>%
  filter(card_type=="X", guess_type=="XandY")
quaternary_summary_XY_and_confint <-
  quaternary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
quaternary_summary_XY_and <- quaternary_summary_XY_and %>% full_join(quaternary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r quaternaryPlot, fig.height=4, fig.cap = "Proportion responses for the four-alternative (quatenary) forced choice judgments in the guessing game."}
quaternary_plot<-
  quaternary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
#  geom_linerange(aes(ymax=cih, ymin=cil), position= position_dodge(width=0.9)) + 
  facet_grid(card_type~guess_type) +
  labs(x="Response", y="Proportion of response")+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values = c("red4", "red3", "springgreen2", "springgreen3"), guide=FALSE) + 
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

Figure \@ref(fig:quaternaryPlot) shows the proportion of "right", "kinda right", "kinda wrong", and "wrong" responses in the quaternary task. Similar to the results seen previously, the control trials turned outas expected. Participants considered a guess "wrong" if the animal guessed was not on the card and "right" if it was the only animal on the card. If a conjunctin of animals was guessed and both animals were on the card the guess was "right". However, when only one of the animals on the card was guessed (exhaustive trials), participants judged the guess "wrong" `r round(quaternaryExh_wrong$proportion*100)`% of the time, "kinda wrong" `r round(quaternaryExh_kindaWrong$proportion*100)`% of the time, and "kinda right" `r round(quaternaryExh_kindaRight$proportion*100)`% of the times. Perhaps surprisingly, when a conjunction was used and only one of the animals was on the card, participants considered the guess "wrong" most of the time, but they often considered it "kinda wrong" or even "kinda right". This suggests that perhaps participants considered a notion of partially true or correct statement in our experimental setting. Disjunctive guesses with one or two animals on the card showed similar response patterns with participants choosing the "kinda right" and "right" options most of the time. When both animals were on the card with a disjunctive guess (scalar trials), participants judged the guess "wrong" `r round(quaternaryScalar_wrong$proportion*100)`% of the time, "kinda wrong" `r round(quaternaryScalar_kindaWrong$proportion*100)`% of the time, and "kinda right" `r round(quaternaryScalar_kindaRight$proportion*100)`% of the times.

@@

quinary_summary_XY_and <-
  quinary_summary %>%
  filter(card_type=="X", guess_type=="XandY")
quinary_summary_XY_and_confint <-
  quinary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
quinary_summary_XY_and <- quinary_summary_XY_and %>% full_join(quinary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r quinaryPlot, fig.height=4, fig.cap = "Proportion responses for the five-alternative (quinary) forced choice judgments in the guessing game."}
quinary_plot<-
  quinary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
#  geom_linerange(aes(ymax=cih, ymin=cil), position= position_dodge(width=0.9)) + 
  facet_grid(card_type~guess_type) +
  labs(x="Response", y="Proportion of response")+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values = c("red4", "red3", "grey", "springgreen2", "springgreen3"), guide=FALSE) +
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

Comparing the response patterns in binary to quinary tasks (Figures \@ref(fig:binaryPlot) to \@ref(fig:quinaryPlot)), we can observe that in implicature trials, participants are less likely to choose the endpoints of the scale (i.e. "wrong" and "right") as they are given more intermediate options. This raises the possibilty that the strict "wrong/right" responses in the tasks with fewer options (binary and ternary) were not accurately reflecting participant judgments of implicature trials and pushed their judgments to the extreme ends. More generally, the data suggest that participant judgments in contexts that give rise to implicatures fall within the intermrediate level of the response scale and forced choice tasks with few options (especially binary tasks) may risk depicting judgments more extreme than they actually are.

@@

Comparing the exhaustive and scalar columns of Figure \@ref(fig:implicatureRatePlot), we see that with a strong linking assumption, there are slightly higher rates for scalar implicatures in the binary and ternary tasks. With a weak linking assumption, there may be slightly higher rates for scalar implicatures in the binary and ternary while the rates may be lower in the quaternary and quinary tasks. In what follows, we test the effect of linking assumption and response options on exhaustive and scalar implicature rates more formally.

```{r implicatureRatePlot, fig.width=5, fig.height=3, fig.cap = "Implicature rate in exhaustive and scalar trials of the binary, ternary, quaternary, and quinary versions of the guessing game, computed once with a strong linking assumption and once with a weak linking assumption."}
ggplot(implicature_results, aes(x=response_type,y=proportion)) +
  geom_bar(aes(fill=definition), stat="identity") +
  facet_grid(definition ~ trial_type, drop = TRUE) +
  theme_few() + 
  labs(x="Implicature") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_linerange(aes(ymin=lower,ymax=upper)) + guides(fill=FALSE)
```

```{r implicatureRate, echo=FALSE, eval=FALSE}
implicature_analysis_a <- glmer(implicature ~ definition + response_type + (1|card) + (1|participant), family="binomial", data=implicature_rate)

summary(implicature_analysis_a)

library(brms)

implicature_analysis_b <- brm(implicature ~ definition * response_type * trial_type + (definition*trial_type*response_type|card) + (definition*trial_type|participant), family="bernoulli", data=implicature_rate, control = list(adapt_delta = 0.99))

saveRDS(implicature_analysis_b, file = "implicature_analysis_b")

summary(implicature_analysis_b)
```

For our formal analysis, we use a binomial mixed effects model with the fixed effect of response type (binary, ternary, quaternary, quinary), linking assumption (strong vs. weak), and trial type (exhaustive vs. scalar), as wellas the random intercept for participants as well as random intercept and slope for items (the card they saw).

All the analytical decisions described here were pre-registered and you can access them through the following link: [https://aspredicted.org/tq3sz.pdf](https://aspredicted.org/tq3sz.pdf)

* Was logical training a significant predictor of "implicature rate"?

# Discussion
=======
On the traditional view of the link between implicature and behavior in sentence verification tasks, scalar implicature is conceptualized as a binary, categorical affair - that is, an implicature is either ‘calculated’ or it isn’t, and the behavioral reflexes of this categorical interpretation process should be straightforwardly observed in experimental paradigms. This assumption has concerning implications for how we must approach analysis of variation in behavior on a truth value judgment task; for example, why did the majority of respondents in the binary condition of our experiment answer “Right” to an utterance of cat or dog when the card had both a cat and a dog on it?

>>>>

```{r load_packages, include = FALSE}
library(grid)
library(gtable)
library(papaja)
library(tidyverse)
library(magrittr)
library(readr)
library(png)
library(jpeg)
library(lme4)
library(ggthemes)
library(forcats)
library(brms)
library(DescTools)
library(binom)
library(knitr)
theme_set(theme_bw(18))
```

@@

Let's examine an example. Assume three response options (wrong, neither, right). Assume further that a third of participants each gave each of the three responses, i.e., the distributions of responses is 1/3, 1/3, and 1/3. Under the Strong link, we infer that this task yielded an implicature rate of 2/3. Under the Weak link, we infer that this task yielded an implicature rate of 1/3. This is quite a drastic difference if we are for instance interested in whether scalar implicatures are inference defaults and we would like to interpret an implicature rate of above an arbitrary threshold (e.g., 50%) as evidence for such a claim. Under the Strong link, we would conclude that scalar implicatures are not defaults. Under the Weak link, we would conclude that they are. In the experiment reported in the following section, we presented participants with exactly this setup.

@@

In this study, we presented participants with an online card game. Different groups of participants were presented with different numbers of response options for the task. In critical trials, participants were presented with descriptions for the cards that typically result in exhaustive or scalar implicatures. We categorized their responses in such trials according to the "weak" and the "strong" link, and tested whether the number of response options and the linking assumptions lead to different conclusions about the rate of computed implicatures in the experimental task.

@@

```{r participants}
participants_info <-
  data %>%
  group_by(response_type) %>%
  summarise(count = n()/24)

mean_age <- mean(data$age)

logic <- data %>% select(participant, logical_training) %>% unique()

logicTraining <- table(logic$logical_training)
```

200 participants were recruited using Amazon Mechanical Turk. They optionally provided demographic information at the end of the study. Mean age of the participnats was `r round(mean_age)`. We also asked participants if they had any prior training in logic. `r logicTraining[[2]]` participants reported that they had while `r logicTraining[[1]]` had no prior training in logic. No participant was excluded from the final analysis.

### Materials and Procedure

The study was administered online through Amazon Mechanical Turk. Participants were first introduced to the set of cards we used in the study (Figure \@ref(fig:stimuli)). Each card had pictures of one or two animals. Animals were chosen from the following set: cat, dog, and elephant. Then participants were introduced to a blindfolded fictional character called Bob. Bob was blindfolded to control for violations of ignorance expectations with disjunction. Participants were told that Bob is going to guess what animals are on the card. We asked participants to let us know whether Bob's guess is wrong or right. In each trial, participants saw a card as well as a sentence representing Bob's guess. For example, they saw a card with a cat and read the sentence "There is a cat on the card." Depending on the task participants were assigned to, they had to choose between two (binary task), three (ternary task), four (quaternary task), or five (quinary task) response options. The study ended after 24 trials. You can access and view the study using the paper's [online repository](https://github.com/thegricean/si-paradigms/).

### Design

@@

The study had two main manipulaitons within participants: the type of card and the type of guess. There were two types of cards. Cards with only one animal on them and cards with two animals. There were three types of guesses: simple (e.g. *There is a cat*), conjunctive (e.g. *There is a cat and a dog*), and disjunctive (e.g. *There is a cat or a dog*). In each trial, the animal labels used in the guess and the animal images on the card may have no overlap (e.g. Image: cat, Guess: *There is an elephant*), a partial overlap (e.g. Image: cat, Guess: *There is a cat or a dog*), or a total overlap (e.g. Image: cat and dog, Guess: *There is a cat or a dog*). Crossing the number of animals on the card, the type of guess, and the overlap between the guess and the card results in 12 different possible trial types. We chose 8 trial types, balancing the number of one-animal vs. two-animal cards, simple vs. connective guesses, and expected true vs. false trials. Three trials were randomly selected from each of the 8 trial-types, for a total of 24 trials. The order of these 24 trials was randomized as well.

@@

The study also had a between participant manipulation of the number of response options in the forced choice task. Participants were randomly assigned to one of four different conditions or tasks. The tasks differed with respect to the number of response options: binary (wrong vs. right), ternary (wrong, neither, right), quaternary (wrong, kinda wrong, kinda right, right), and quinary (wrong, kinda wrong, neither, kinda right, right). We wanted to see if the number of response options in the forced choice task would affect our estimate of the task's "implicature rate".

@@

@@

binary_summary_XY_and <-
  binary_summary %>%
  filter(card_type=="XY", guess_type=="XandY")
binary_summary_XY_and_confint <-
  binary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
binary_summary_XY_and <- binary_summary_XY_and %>% full_join(binary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r binaryPlot, fig.height=3.5, fig.cap = "Proportion responses for the two-alternative (binary) forced choice judgments in the guessing game."}
binary_plot<-
  binary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_grid(card_type~guess_type) +
  labs(x="Response Options in the Binary Task", y="Proportions")+
  theme_few() +
  theme(text = element_text(size=12)) +
  scale_fill_manual(values = c("red4", "springgreen3"), guide=FALSE) +
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

The experiment had `r participants_info$count[1]` participants in the binary task, `r participants_info$count[4]` in the ternary task, `r participants_info$count[2]` in the quaternary task and `r participants_info$count[3]` in the quinary task. In this section, we present the proportion that participants chose different response options in each of the 8 trial types of these four tasks.

Figure \@ref(fig:binaryPlot) shows the proportion of "right" and "wrong" responses in the binary task. Starting from the leftmost column, participants considered a guess "wrong" if the guessed animal was not on the card. Moving to the second column, participants considered the guess "right" when the only animal on the card was mentioned. However, when only one of the two animals on the card was mentioned (exhaustive trials), `r round(binaryExh_wrong$proportion*100)`% of the times participants considered the guess "wrong". Moving to the third column, if a conjunction of animals was guessed while only one animal was on the card, participants considered the guess to be "wrong". If a conjunction of animals was guessed and both animals were present on the card, all participants considered the guess to be "right" as expected. Finally, if a disjunction of animals was guessed and only one of the animals was on the card, participants considered the guess to be "right" almost all the time. However, if both animals were present (scalar trials), `r round(binaryScalar_wrong$proportion*100)`% of the times participants considered the guess to be "wrong".

@@

ternary_summary_XY_and <-
  ternary_summary %>%
  filter(card_type=="XY", guess_type=="XandY")
ternary_summary_XY_and_confint <-
  ternary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
ternary_summary_XY_and <- ternary_summary_XY_and %>% full_join(ternary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r ternaryPlot, fig.height=3.5, fig.cap = "Proportion responses for the three-alternative (ternary) forced choice judgments in the guessing game."}
ternary_plot<-
  ternary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
  facet_grid(card_type~guess_type) +
  labs(x="Response Options in the Ternary Task", y="Proportions")+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = c("red4","grey", "springgreen3"), guide=FALSE) +
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

Figure \@ref(fig:ternaryPlot) shows the proportion of "right", "neither", and "wrong" responses in the ternary task. Similar to the binary task, participants considered a guess wrong when the mentioned animal was not on the card, and "right" when the mentioned animal was the only animal on the card. However, in exhaustive trials when the fictional character only guessed one of the two animals on the card, participants considered the guess "wrong" `r round(ternaryExh_wrong$proportion*100)`% of the time and neither wrong nor right `r round(ternaryExh_neither$proportion*100)`% of the time. If a conjunction of animals was guessed and only one animal was present on the card, participants considered the guess "wrong". As expected, when a conjunction was used and both animals were present, participants considered the guess "right". Similarly, participants considerd the guess "right" when a disjunction was used and only one of the animals was on the card. However, in scalar trials that both animals were on the card and a disjunction was gueassed, participants judged the guess "wrong" `r round(ternaryScalar_wrong$proportion*100)`% of the time and "neither" `r round(ternaryScalar_neither$proportion*100)`% of the time.

@@

quaternary_summary_XY_and <-
  quaternary_summary %>%
  filter(card_type=="XY", guess_type=="XandY")
quaternary_summary_XY_and_confint <-
  quaternary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
quaternary_summary_XY_and <- quaternary_summary_XY_and %>% full_join(quaternary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r quaternaryPlot, fig.height=4, fig.cap = "Proportion responses for the four-alternative (quatenary) forced choice judgments in the guessing game."}
quaternary_plot<-
  quaternary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
#  geom_linerange(aes(ymax=cih, ymin=cil), position= position_dodge(width=0.9)) + 
  facet_grid(card_type~guess_type) +
  labs(x="Response Options for the Quaternary Task", y="Proportions")+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values = c("red4", "red3", "springgreen2", "springgreen3"), guide=FALSE) + 
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

Figure \@ref(fig:quaternaryPlot) shows the proportion of "right", "kinda right", "kinda wrong", and "wrong" responses in the quaternary task. Similar to the results seen previously, the control trials turned out as expected. Participants considered a guess "wrong" if the animal guessed was not on the card and "right" if it was the only animal on the card. If a conjunctin of animals was guessed and both animals were on the card the guess was "right". However, when only one of the animals on the card was guessed (exhaustive trials), participants judged the guess "wrong" `r round(quaternaryExh_wrong$proportion*100)`% of the time, "kinda wrong" `r round(quaternaryExh_kindaWrong$proportion*100)`% of the time, and "kinda right" `r round(quaternaryExh_kindaRight$proportion*100)`% of the times. Perhaps surprisingly, when a conjunction was used and only one of the animals was on the card, participants considered the guess "wrong" most of the time, but they often considered it "kinda wrong" or even "kinda right". This suggests that perhaps participants considered a notion of partially true or correct statement in our experimental setting. Disjunctive guesses with one or two animals on the card showed similar response patterns with participants choosing the "kinda right" and "right" options most of the time. When both animals were on the card with a disjunctive guess (scalar trials), participants judged the guess "wrong" `r round(quaternaryScalar_wrong$proportion*100)`% of the time, "kinda wrong" `r round(quaternaryScalar_kindaWrong$proportion*100)`% of the time, and "kinda right" `r round(quaternaryScalar_kindaRight$proportion*100)`% of the times.

@@

quinary_summary_XY_and <-
  quinary_summary %>%
  filter(card_type=="XY", guess_type=="XandY")
quinary_summary_XY_and_confint <-
  quinary_summary_XY_and$count %>%
  MultinomCI(conf.level = 0.95, method = "sisonglaz") %>% data.frame()
quinary_summary_XY_and <- quinary_summary_XY_and %>% full_join(quinary_summary_XY_and_confint, by="est") %>% unique()

@@

```{r quinaryPlot, fig.height=4, fig.cap = "Proportion responses for the five-alternative (quinary) forced choice judgments in the guessing game."}
quinary_plot<-
  quinary_summary %>%
  ggplot(aes(x=response, y=proportion, fill=response)) +
  geom_bar(stat = "identity", position="dodge", width = 0.6) +
#  geom_linerange(aes(ymax=cih, ymin=cil), position= position_dodge(width=0.9)) + 
  facet_grid(card_type~guess_type) +
  labs(x="Response Options in the Quinary Task", y="Proportions")+
  theme_few() +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_fill_manual(values = c("red4", "red3", "grey", "springgreen2", "springgreen3"), guide=FALSE) +
  geom_linerange(aes(ymin=lwr.ci,ymax=upr.ci))

@@

<!--
In four trial types, participants provided the same responses regardless of how many response options were made available. The first two trial types include the guesses with an animal not on the card at all. In such trials, whether there was one or two animals on the card, participants considered the guess wrong. In the third and the fourth trial types, the gueses matched the aimal(s) on the card; if there was one animal on the card, that animal was mentioned and if there were two, both were mentioned using a conjunction. The fact that participants chose the extreme ends of the scale for these trial types and the responses did not change when intermediate options were provided suggests that participant judgments truly aligned with the opposite ends of the scale in these trial types.

In the trial type where the guess was a conjunction but only one of the animals was on the card, participants considered the guess wrong when fewer options were provided. However, in tasks with more options (four and five), participant responses were distributed between "wrong", "kinda wrong", and "kinda right" options, with the majority of responses being "wrong" or "kinda wrong". The graded nature of the judgments in this trial type suggests that participants do not consider such guesses as completely wrong. Similarly,

Finally, in exhaustive and scalar implicature trial types participants were less likely to choose the option "wrong" when they were given more intermediate options. Even though in the binary task task recorded participant judgments as

semantic violations vs. pragmatic violations
-->

@@

Comparing the exhaustive and scalar columns of Figure \@ref(fig:implicatureRatePlot), we see that with a strong linking assumption, there are slightly higher rates for scalar implicatures in the binary and ternary tasks. With a weak linking assumption, there may be slightly higher rates for scalar implicatures in the binary and ternary while the rates may be lower in the quaternary and quinary tasks. In what follows, we formally test the effect of linking assumption and response options on exhaustive and scalar implicature rates.

```{r implicatureRatePlot, fig.width=5, fig.height=3, fig.cap = "Implicature rate in exhaustive and scalar trials of the binary, ternary, quaternary, and quinary versions of the guessing game, computed once with a strong linking assumption and once with a weak linking assumption."}
ggplot(implicature_results, aes(x=response_type,y=proportion)) +
  geom_bar(aes(fill=definition), stat="identity") +
  facet_grid(trial_type~definition, drop = TRUE) +
  theme_few() + 
  labs(x="Number of Response Options in the Task", y="Inferred Implicature Rate") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_linerange(aes(ymin=lower,ymax=upper)) + guides(fill=FALSE)
```

```{r implicatureRate, warning=FALSE}
# The analysis takes time to run. To reproduce please run the commented lines. We have saved the results we obtained as a dataframe.

#library(brms)

#implicature_analysis_b <- brm(implicature ~ definition * response_type * trial_type + (definition*trial_type*response_type|card) + (definition*trial_type|participant), family="bernoulli", data=implicature_rate, control = list(adapt_delta = 0.99, max_treedepth = 15))

#saveRDS(implicature_analysis_b, file = "implicature_analysis_b")
#implicature_analysis_b <- readRDS("implicature_analysis_b")

#fixed_effects <- data.frame(fixef(implicature_analysis_b))
#write.csv(fixed_effects, "experiments/main/fixed_effect_estimates.csv")

fixed_effects <- read_csv("experiments/main/fixed_effect_estimates.csv")

fixed_effects$X1 <- c("Intercept", "Link = Weak", "Task = Quaternary", "Task = Quinary", "Task = Ternary", "Implicature = Scalar", "Link = Weak : Task = Quaternary", "Link = Weak : Task = Quinary", "Link = Weak : Task = Ternary", "Link = Weak : Implicature = Scalar", "Task = Quaternary : Implicature = Scalar", "Task = Quinary : Implicature = Scalar", "Task = Ternary : Implicature = Scalar", "Link=Weak : Task=Quaternary : Implicature=Scalar", "Link=Weak : Task=Quinary : Implicature=Scalar", "Link=Weak : Task=Ternary : Implicature=Scalar")

fixed_effects_table <- 
  fixed_effects %>%
  rename(Predictors = "X1", `2.5%` = "X2.5.ile", `97.5%` = "X97.5.ile") %>%
  select(-Est.Error)

kable(fixed_effects_table,digits = 2 , caption = "\\label{tab:modeltable}Model Parameter Estimates and Their Credible Intervals", booktabs = T)
```

For our formal analysis, we used a bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms]. The model had the fixed effects of response type (binary, ternary, quaternary, quinary), linking assumption (strong vs. weak), and trial type (exhaustive vs. scalar), as well as their two way and three way interactions. We also included random intercepts and slopes for items (cards) and participants. Following @barr2013random, we used the maximal random effects structure by including random slopes for all our fixed effects and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^1].

The model provided evidence for two effects in the study. First, there was a main effect of trials such that scalar trials had slightly higher implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between linking assumption and number of resopnse options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a weak linking assumption had higher implicature rates.

[^1]: You can access our pre-registration at [https://aspredicted.org/tq3sz.pdf](https://aspredicted.org/tq3sz.pdf)

## Discussion

We asked whether the linking assumptions and the number of response options available to participants affects the inferred implicature rate in an experimental study. The results presented here suggest they do. A linking assupmtion that considered the highest point on the scale as literal and any lower point as pragmatic (weak link) reported higher implicature rates in tasks with 4 or 5 options. A linking assumption that considered the lowest point on the scale as pragmatic and any higher point as literal (strong link) reported lower implicature rates in tasks with 4 or 5 options. The results suggest that the choice of linking assumption is a crucial analytical step that can significantly impact the conclusions of truth value judgment tasks with more than two response options. The lower rate of implicatures with a strong linking assumption implies that in such studies, strong linking assumptions may underestimate participants' pragmatic competence.

While the binary truth value judgement task avoids the analytic decision between strong and weak linking assumptions, our results suggest that binary tasks can also underestimate participants' pragmatic competence. In binary tasks, participants are often given the lowest and highest points on a scale ("wrong" vs. "right") and are asked to report pragmatic infelicities using the lowest point (e.g. "wrong"). Our study showed that in trials with true but infelicitous descriptions (implicature trials), participants often avoided the lowest point on the scale if they were given more intermediate options. Even though the option "wrong" was available to participants in all tasks, participants in tasks with intermediate options chose it less often. In computing implicature rate, this pattern manifested itself as a decrease in implicature rate with strong link when more response options were provided, and increase in implicature rate with weak link when more response options were provided. These conclusions are in line with @Katsos2011's argument that pragmatic violations are not as severe as semantic violations and participants do not penalize them as much. Providing participants with only the extreme opposits of the scale (e.g. wrong/right, false/true) when pragmatic violations are considered to be of an intermediate nature risks misrepresentation of participants' pragmatic competence.

This study did not investigate the effect of option labels on the inferred implicature rate. However, the results provided suggestive evidence that some options better capture participant intuitions of pragmatic infelicities than others. Among the intermediate options, "kinda right" was chosen most often to report pragmatic infelicities. The option "neither" was rarely used in the ternary and quinary tasks (where it was used as a midpoint), suggesting that participants found pragmatic infelicities as different degrees of being "right" and not "neither right nor wrong." Therefore, options that capture grades of being "right" like "kinda right" proved be most suitable for capturing true but infelicitous utterances.
